---
title: "2026 FIFA World Cup Players Age Distribution"
author: "Anthony Gao, Antigravity"
date: "`r Sys.Date()`"
format: html
---

## Introduction

This report analyzes the age distribution of the players participating in the 2026 FIFA World Cup, utilizing data scraped from Wikipedia.

```{r setup}
#| include: false
#| message: false
#| warning: false
suppressPackageStartupMessages(library(tidyverse))

# Read the players dataset
players <- read_csv("players.csv")
```



## Age Distribution by Position

Below is the age distribution comparison across different player positions represented as overlaid density curves:

```{r density-position}
#| echo: false
#| message: false
#| warning: false
ggplot(players, aes(x = Age, fill = Position, color = Position)) +
  geom_density(alpha = 0.2, linewidth = 1) +
  scale_fill_manual(values = c("GK" = "#e63946", "DF" = "#457b9d", "MF" = "#1d3557", "FW" = "#2a9d8f")) +
  scale_color_manual(values = c("GK" = "#e63946", "DF" = "#457b9d", "MF" = "#1d3557", "FW" = "#2a9d8f")) +
  scale_x_continuous(breaks = seq(min(players$Age), max(players$Age), by = 2)) +
  labs(
    title = "Age Density Distribution by Position",
    x = "Age",
    y = "Density",
    fill = "Position",
    color = "Position"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )
```



## Average Age of Squads by Team (Grouped by Continent)

Below is a lollipop chart displaying the average age of the squads for each team, sorted by GDP:

```{r average-age-team}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 12

# Named vector mapping each country to its respective continent and dummy GDP data
continent_mapping <- c(
  "Czech Republic" = "Europe", "Mexico" = "North America", "South Africa" = "Africa",
  "South Korea" = "Asia", "Bosnia and Herzegovina" = "Europe", "Canada" = "North America",
  "Qatar" = "Asia", "Switzerland" = "Europe", "Brazil" = "South America",
  "Haiti" = "North America", "Morocco" = "Africa", "Scotland" = "Europe",
  "Australia" = "Asia", "Paraguay" = "South America", "Turkey" = "Europe",
  "United States" = "North America", "Curaçao" = "North America", "Ecuador" = "South America",
  "Germany" = "Europe", "Ivory Coast" = "Africa", "Japan" = "Asia",
  "Netherlands" = "Europe", "Sweden" = "Europe", "Tunisia" = "Africa",
  "Belgium" = "Europe", "Egypt" = "Africa", "Iran" = "Asia",
  "New Zealand" = "Oceania", "Cape Verde" = "Africa", "Saudi Arabia" = "Asia",
  "Spain" = "Europe", "Uruguay" = "South America", "France" = "Europe",
  "Iraq" = "Asia", "Norway" = "Europe", "Senegal" = "Africa",
  "Algeria" = "Africa", "Argentina" = "South America", "Austria" = "Europe",
  "Jordan" = "Asia", "Colombia" = "South America", "DR Congo" = "Africa",
  "Portugal" = "Europe", "Uzbekistan" = "Asia", "Croatia" = "Europe",
  "England" = "Europe", "Ghana" = "Africa", "Panama" = "North America"
)

# GDP (nominal, USD billions, approx 2023/2024 IMF/World Bank figures)
gdp_data <- c(
  "United States" = 27360, "Germany" = 4457, "Japan" = 4213,
  "France" = 3031, "England" = 3100, "Brazil" = 2174,
  "Canada" = 2140, "Australia" = 1723, "Mexico" = 1788,
  "South Korea" = 1712, "Spain" = 1582, "Saudi Arabia" = 1062,
  "Netherlands" = 1118, "Turkey" = 1108, "Switzerland" = 905,
  "Sweden" = 597, "Belgium" = 627, "Argentina" = 621,
  "Norway" = 546, "Austria" = 527, "Iran" = 413,
  "South Africa" = 377, "Egypt" = 395, "Colombia" = 363,
  "Czech Republic" = 341, "Scotland" = 250, "New Zealand" = 249,
  "Algeria" = 239, "Qatar" = 235, "Iraq" = 264,
  "Portugal" = 287, "Morocco" = 143, "Ecuador" = 118,
  "Uzbekistan" = 90, "Uruguay" = 78, "Panama" = 77,
  "Ghana" = 77, "Ivory Coast" = 71, "DR Congo" = 65,
  "Croatia" = 82, "Tunisia" = 49, "Jordan" = 50,
  "Paraguay" = 43, "Senegal" = 27, "Bosnia and Herzegovina" = 25,
  "Haiti" = 20, "Cape Verde" = 2, "Curaçao" = 4
)

team_data <- players %>%
  group_by(Country) %>%
  summarise(
    AverageAge = mean(Age),
    .groups = "drop"
  ) %>%
  mutate(
    Continent = continent_mapping[Country],
    GDP_bn = gdp_data[Country]
  ) %>%
  arrange(desc(GDP_bn)) %>%
  mutate(Country = factor(Country, levels = rev(Country)))

ggplot(team_data, aes(x = AverageAge, y = Country, color = Continent)) +
  geom_segment(aes(x = 22, xend = AverageAge, y = Country, yend = Country), color = "grey80", linewidth = 0.5) +
  geom_point(size = 2.5) +
  geom_text(aes(label = sprintf("%.1f", AverageAge)), hjust = -0.4, size = 2.5, show.legend = FALSE) +
  xlim(22, 31) +
  scale_color_manual(values = c(
    "Africa" = "#e63946",
    "Asia" = "#2a9d8f",
    "Europe" = "#1d3557",
    "North America" = "#f4a261",
    "South America" = "#e76f51",
    "Oceania" = "#9b5de5"
  )) +
  labs(
    title = "Average Age of Squads by Team (Ordered by GDP, Highest to Lowest)",
    x = "Average Age (Years)",
    y = "Team",
    color = "Continent"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )
```

## Average Age vs. Country GDP (by Continent)

The scatter plot below shows the relationship between a country's nominal GDP and its squad's average age. Each point represents one national team, colored by continent. A LOESS trend line is overlaid to highlight any overall pattern.

```{r age-vs-gdp-scatter}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 7

continent_mapping <- c(
  "Czech Republic" = "Europe", "Mexico" = "North America", "South Africa" = "Africa",
  "South Korea" = "Asia", "Bosnia and Herzegovina" = "Europe", "Canada" = "North America",
  "Qatar" = "Asia", "Switzerland" = "Europe", "Brazil" = "South America",
  "Haiti" = "North America", "Morocco" = "Africa", "Scotland" = "Europe",
  "Australia" = "Asia", "Paraguay" = "South America", "Turkey" = "Europe",
  "United States" = "North America", "Curaçao" = "North America", "Ecuador" = "South America",
  "Germany" = "Europe", "Ivory Coast" = "Africa", "Japan" = "Asia",
  "Netherlands" = "Europe", "Sweden" = "Europe", "Tunisia" = "Africa",
  "Belgium" = "Europe", "Egypt" = "Africa", "Iran" = "Asia",
  "New Zealand" = "Oceania", "Cape Verde" = "Africa", "Saudi Arabia" = "Asia",
  "Spain" = "Europe", "Uruguay" = "South America", "France" = "Europe",
  "Iraq" = "Asia", "Norway" = "Europe", "Senegal" = "Africa",
  "Algeria" = "Africa", "Argentina" = "South America", "Austria" = "Europe",
  "Jordan" = "Asia", "Colombia" = "South America", "DR Congo" = "Africa",
  "Portugal" = "Europe", "Uzbekistan" = "Asia", "Croatia" = "Europe",
  "England" = "Europe", "Ghana" = "Africa", "Panama" = "North America"
)

gdp_data <- c(
  "United States" = 27360, "Germany" = 4457, "Japan" = 4213,
  "France" = 3031, "England" = 3100, "Brazil" = 2174,
  "Canada" = 2140, "Australia" = 1723, "Mexico" = 1788,
  "South Korea" = 1712, "Spain" = 1582, "Saudi Arabia" = 1062,
  "Netherlands" = 1118, "Turkey" = 1108, "Switzerland" = 905,
  "Sweden" = 597, "Belgium" = 627, "Argentina" = 621,
  "Norway" = 546, "Austria" = 527, "Iran" = 413,
  "South Africa" = 377, "Egypt" = 395, "Colombia" = 363,
  "Czech Republic" = 341, "Scotland" = 250, "New Zealand" = 249,
  "Algeria" = 239, "Qatar" = 235, "Iraq" = 264,
  "Portugal" = 287, "Morocco" = 143, "Ecuador" = 118,
  "Uzbekistan" = 90, "Uruguay" = 78, "Panama" = 77,
  "Ghana" = 77, "Ivory Coast" = 71, "DR Congo" = 65,
  "Croatia" = 82, "Tunisia" = 49, "Jordan" = 50,
  "Paraguay" = 43, "Senegal" = 27, "Bosnia and Herzegovina" = 25,
  "Haiti" = 20, "Cape Verde" = 2, "Curaçao" = 4
)

scatter_data <- players %>%
  group_by(Country) %>%
  summarise(AverageAge = mean(Age), .groups = "drop") %>%
  mutate(
    Continent = continent_mapping[Country],
    GDP_bn    = gdp_data[Country]
  ) %>%
  filter(!is.na(GDP_bn), !is.na(Continent))

continent_colors <- c(
  "Africa"        = "#e63946",
  "Asia"          = "#2a9d8f",
  "Europe"        = "#1d3557",
  "North America" = "#f4a261",
  "South America" = "#e76f51",
  "Oceania"       = "#9b5de5"
)

ggplot(scatter_data, aes(x = GDP_bn, y = AverageAge, color = Continent, fill = Continent)) +
  geom_smooth(
    aes(group = Continent),
    method    = "loess",
    formula   = y ~ x,
    se        = FALSE,
    linewidth = 0.8
  ) +
  geom_point(size = 3.5, alpha = 0.9) +
  ggrepel::geom_text_repel(
    aes(label = Country),
    size          = 2.8,
    max.overlaps  = 20,
    segment.color = "grey60",
    segment.size  = 0.3,
    show.legend   = FALSE
  ) +
  scale_x_log10(
    breaks = c(2, 10, 50, 250, 1000, 5000, 25000),
    labels = scales::label_comma(suffix = "B")
  ) +
  scale_color_manual(values = continent_colors) +
  scale_fill_manual(values = continent_colors) +
  labs(
    title    = "Average Squad Age vs. Country GDP",
    subtitle = "Each point is one national team · GDP on a log scale · Per-continent LOESS trend lines shown",
    x        = "Nominal GDP (USD billions, log scale)",
    y        = "Average Squad Age (Years)",
    color    = "Continent",
    fill     = "Continent"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title    = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, color = "grey40", size = 10),
    panel.grid.minor  = element_blank(),
    legend.position   = "bottom",
    legend.title      = element_text(face = "bold")
  )
```

## Summary Statistics

Below are the summary statistics of the players' ages:


```{r summary}
#| echo: false
players %>%
  summarise(
    `Minimum Age` = min(Age),
    `First Quartile` = quantile(Age, 0.25),
    `Median Age` = median(Age),
    `Average Age` = round(mean(Age), 2),
    `Third Quartile` = quantile(Age, 0.75),
    `Maximum Age` = max(Age)
  ) %>%
  knitr::kable(align = 'c')
```

